---
title: "Implementation"
output: rmarkdown::html_vignette
bibliography: Ref.bib
vignette: >
  %\VignetteIndexEntry{Implementation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r,echo=FALSE}
knitr::read_chunk("../R/DMIX.R")
```


```{r setup}
library(DMIX)
```


The `DMIX` package fits Normal mixtures to binned length density data from fish and plankton surveys.
The models are similar to those proposed by [@DLM94CMIX], but are fitted by Bayesian sampling 
techniques [@Plummer2003].


## Delta Log Normal Densities

Following [@Pennington83] the delta log Normal model assumes that the density in each length bin has
a delta log Normal distribution.

The observed density $x_{ki}$ in length bin $[l_{i}, l_{i+1})$ in sample $k$ is assume to be delta
log Normal
$$
y_{ki} = 
\begin{cases}
0 & z_{ki} = 0\\
x_{ki} & z_{ki} = 1
\end{cases}
$$
where
$$
\begin{align}
z_{ki} &\sim \operatorname{Bin}(1,p_{i})\\
\log x_{ki} &\sim \operatorname{N}(m_{i},s_{i}^{2}) 
\end{align}
$$
and $p_{i}$ is the probability of observing a non-zero density in length bin $i$, and $m_{i}$ and 
$s_{i}^{2}$ are the mean and variance of the log of the non-zero observations in length bin $i$.  
It follows that the mean density $\lambda_{i} = \operatorname{E}(y_{ki})$ in each bin is 
$$
\lambda_{i} = p_{i} \exp(m_{i} + s_{i}^{2}/2).
$$
In turn the mean density in each bin is assumed to be determined by a Normal mixture integrated over 
the length bin
$$
\lambda_{i} = \sum_{k=1}^{m} w_{k} [\Phi(l_{i-1}; \mu_{k}, \sigma_{k}^2) -\Phi(l_{i}; \mu_{k}, \sigma_{k}^2)]
$$
where $w_{k}$ is the total density in component $k$, $\mu_{k}$ and $\sigma_{k}^{2}$ are the mean and 
variance of component $k$, and  $\Phi(l; \mu, \sigma)$ is the distribution function of the Normal.

The probability $p_{i}$ of a non-zero density and the dispersion $s_{i}$ across the bins are modelled 
as either

* as fixed effects with $p_{i}$ and $s_{i}$ estimated in each bin
* constant across bins, $p_{i} = p$ and $s_{i}=s$ or
* as independent Normal random effects
$$
\begin{align}
\operatorname{logit} p_{i} &\sim \operatorname{N}(\mu_{p},\sigma_{p}^{2})\\
\log s_{i} &\sim \operatorname{N}(\mu_{s},\sigma_{s}^{2})
\end{align}
$$


Similarly, there are three alternative models for the standard deviation of the mixture components

* the $\sigma_{k}$ are estimated independently for each component 
* the $\sigma_{k} = \kappa \mu_{k}$ are assumed proportional to the $\mu_{k}$
* the $\sigma_{k} = a + b \mu_{k}$ are assumed to be linearly related to the $\mu_{k}$.


## Reference

